---
title: "Final Course Project"
subtitle: "Final Project in R Markdown"
author: "*Kurtik Appadoo and Kyle Zaslaw*"
output: 
  pdf_document:
    toc: yes
    number_sections: yes
date: "`r format(Sys.Date(), '%B %Y')`"  
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include=FALSE}
# setting chunk options globally so that you don't have to specify each time.
knitr::opts_chunk$set(out.width='50%', fig.asp=.7, fig.path='../output/figures/', fig.align = 'center', warning=FALSE, message=FALSE, collapse = TRUE, echo = FALSE)

# sets the number of digits
options(digits=4)
```

```{r, include=FALSE}
# activate all packages that you will need for sure, for example 
library(tinytex) # required to produce pdf documents
library(tidyverse)
library(fpp3)
library(forecast)
```

```{r}
data <- read_csv("data1.csv") |> 
                mutate(DATE = yearquarter(DATE))
data_long <- read_csv("data_long1.csv") |> 
                mutate(DATE = yearquarter(DATE))
data_ts <- data |> 
  as_tsibble(index = DATE)
data_long_ts <- data_long |> 
  as_tsibble(index = DATE, key = Series)
```

```{r}
data2 <- read_csv("data2.csv") |> 
                mutate(DATE = yearmonth(DATE))
data_long2 <- read_csv("data_long2.csv") |> 
                mutate(DATE = yearmonth(DATE))
data_ts2 <- data2 |> 
  as_tsibble(index = DATE)
data_long_ts2 <- data_long2 |> 
  as_tsibble(index = DATE, key = Series)
```

```{r}
data3 <- read_csv("data3.csv") |> 
                mutate(DATE = yearquarter(DATE))
data_long3 <- read_csv("data_long3.csv") |> 
                mutate(DATE = yearquarter(DATE))
data_ts3 <- data3 |> 
  as_tsibble(index = DATE)
data_long_ts3 <- data_long3 |> 
  as_tsibble(index = DATE, key = Series)
```

```{r, include=FALSE}
data4 <- read_csv("data4.csv") |> 
                mutate(DATE = yearmonth(DATE))
data_long4 <- read_csv("data_long4.csv") |> 
                mutate(DATE = yearmonth(DATE))
data_ts4 <- data4 |> 
  as_tsibble(index = DATE)
data_long_ts4 <- data_long4 |> 
  as_tsibble(index = DATE, key = Series)
```

\newpage

# Introduction

Forecasting key economic indicators such as Real GDP Growth, CPI
(Consumer Price Index), Real Non-Residential Fixed Investment, and
Average Weekly Earnings is pivotal for comprehending and guiding the
economic trajectory of a country. These forecasts serve as the backbone
for economic policy-making, investment decisions, and provide insights
into the overall health and direction of the economy. Forecasting these
economic indicators is critical for several intertwined reasons.
Understanding CPI in conjunction with Real GDP Growth helps central
banks and policymakers balance growth with inflation control, crucial
for maintaining purchasing power and economic stability. Forecasting
Real Non-Residential Fixed Investment sheds light on business sentiment
and future productivity, as these investments are directly tied to
capacity expansion and technological advancements, driving long-term GDP
growth. Average Weekly Earnings offer insights into consumer spending
potential, inflationary pressures, and the health of the labor market,
which in turn influences Real GDP Growth through consumer spending.

To holistically forecast these indicators, we'll employ a few
forecasting models we've studied in class as well as contribute the
effect of various predictors in helping the accuracy of our model and
thus forecast. Given the time series nature of these indicators, ARIMA
models, augmented with seasonal adjustments, will capture underlying
trends, cycles, and seasonal effects, crucial for accurate short-term
and medium-term forecasts. To avoid using just one model, we'll also
establish and compare Time series Linear models to encapsulate the
underlying and existing linear relationships in existing economic
models. As we tend to observe in most economic literature, ARIMA models
or just models that make use of lagged variables tend to yield more
accurate results, however some TSLM models have proven to provide useful
insights, Thus we'll set up and compare both.

Accurate forecasting and analysis could unveil; How inflation dynamics
(CPI) interact with GDP growth, the implications of fixed investment on
future productivity and GDP, and the feedback loop between earnings
growth, consumer spending, and overall economic activity. Evaluate the
effectiveness of monetary and fiscal policies on stabilizing inflation,
stimulating investment, and enhancing workers' earnings without
overheating the economy. Identifying emerging trends and potential risks
in inflation, investment, and labor markets, guiding policymakers and
investors in proactive decision-making. Utilizing forecasts to fine-tune
monetary policy, ensuring sustainable growth while keeping inflation in
check, thus preserving the economy's purchasing power. Fiscal incentives
or support measures could be designed to encourage non-residential fixed
investment in sectors identified as growth drivers, enhancing long-term
economic prospects. Understanding the trajectory of average weekly
earnings alongside labor market trends can inform minimum wage
adjustments, tax policies, and social welfare programs to support
equitable growth.

In blending these forecasts, we gain a comprehensive view of the
economy, allowing for nuanced policy interventions that support
sustainable growth, manage inflation, encourage productive investment,
and ensure the well-being of the workforce. This integrated approach
enhances our ability to navigate economic complexities, leverage
opportunities for growth, and mitigate potential risks.

# Literature Review

### Real GDP growth quarterly

Consumer Expenditure Indicates consumer confidence and disposable
income, driving demand and GDP growth. Lagged variable: Identifies
trends in consumer behavior over time, affecting future economic
activity.

Business Investment Reflects business confidence and is crucial for
productivity and expanding production capacity. Lagged variable: Gauges
the delayed impact of investments on economic growth, considering the
time to fruition.

Government Spending/Investment Directly boosts GDP through public
services and infrastructure, stimulating economic activity. Lagged
variable: Helps understand the rollout and impact of fiscal policies on
future GDP growth.

Interest Rates Influence borrowing costs, spending, investment, and
thereby economic growth. Affect exchange rates and export
competitiveness. Lagged variable: Accounts for the time lag in monetary
policy's effect on the economy, guiding future GDP predictions.

Lagged data for these predictors is crucial in forecasting real GDP
growth, as it captures the time-delayed effects of economic activities
and policies on the overall economy, enhancing forecast accuracy.

### Average Weekly Earnings monthly

Employment Rate Reflects labor demand; higher employment rates can lead
to wage increases due to competition for labor. Lagged Data: Wage
adjustments may follow changes in employment rates as employers react to
shifting labor market conditions over time.

Inflation Rate (CPI) Workers seek wages that keep pace with the cost of
living, influencing wage negotiations. Lagged Data: Past inflation
trends help predict wage adjustments as both employers and employees
consider inflation in their wage determinations.

Labor Productivity Higher productivity can lead to wage increases as
businesses generate more revenue per employee. Lagged Data: Productivity
gains often translate into wage growth after businesses assess the
durability of these improvements.

Industry Growth Rates Fast-growing industries may offer higher wages due
to increased demand for skilled labor. Lagged Data: Wages adjust
following industry growth, reflecting a period of evaluation and
financial planning by businesses.

Understanding these predictors and incorporating lagged data allows for
a more accurate forecast of average weekly earnings, accounting for the
time it takes for economic and policy changes to fully impact wage
levels.

### CPI monthly (index)

Producer Price Index (PPI): The PPI measures the average changes in
selling prices received by domestic producers for their output and is
often seen as a leading indicator for CPI. Increases in PPI can indicate
rising costs for producers that may be passed on to consumers, leading
to higher CPI.

Lagged Data Usage: A short lag, often one to two months, is useful since
changes in producer prices quickly translate to consumer prices,
especially in fast-moving consumer goods.

Wage Growth: Rising wages can increase disposable income, leading to
higher demand for goods and services, which can push prices up. Wage
growth is a component of cost-push inflation, where the cost of
production increases, leading to higher prices.

Lagged Data Usage: A lag of a few months is often considered, as changes
in wage policies or labor market conditions may take time to manifest in
consumer prices.

Oil Prices: Oil prices directly affect the cost of transportation and
production of goods. High oil prices can lead to increased production
costs for goods and services, contributing to higher CPI. Lagged Data
Usage: The effect of oil price changes on CPI can be immediate or
delayed, depending on the extent to which businesses absorb the
increased costs before passing them on to consumers, usually considering
a lag of one to three months.

Exchange Rates: The strength of a country's currency can influence
inflation through import prices. A weaker currency makes imports more
expensive, contributing to higher consumer prices, whereas a stronger
currency can have the opposite effect. Lagged Data Usage: Exchange rate
movements may take several months to influence CPI as changes in import
costs are gradually passed on to consumer prices.

Economic Activity Indicators (e.g., GDP Growth, Unemployment Rate): High
levels of economic activity and low unemployment can lead to increased
demand for goods and services, potentially causing prices to rise if
supply does not keep pace with demand.

Lagged Data Usage: These indicators may have varying lags, with GDP data
often considered on a quarterly basis and unemployment rates on a
monthly basis. The impact on CPI may be observed over several months as
changes in economic activity affect consumer spending patterns.

### Real non-Residential Fixed Investments growth quarterly

Interest Rates: Interest rates are a critical factor in investment
decisions. Lower interest rates reduce the cost of borrowing, making it
cheaper for businesses to finance new projects and investments, whereas
higher rates do the opposite. Lagged Data Usage: The effect of interest
rate changes on investment can have a lag, as firms adjust their
investment plans and financing arrangements. A lag of a few months to a
year can be considered.

Corporate Profits: Higher corporate profits can signal strong business
conditions, providing firms with the internal funds needed for
investment. It indicates the financial health and potential for
reinvestment into productive assets. Lagged Data Usage: Investments
driven by corporate profits might be observed with a lag, as decisions
on allocating profits into new investments are made over time, typically
considering a lag of one to two quarters.

Technological Changes and Innovations: Technological advancements can
create new investment opportunities by improving the efficiency or
reducing the cost of production. Businesses need to invest in new
technology to remain competitive and capitalize on these advancements.
Lagged Data Usage: The adoption of new technologies can have a varied
lag, as firms need time to assess, plan, and implement technological
investments, often over months to several years depending on the sector
and technology complexity.

Depending on the number of authors, use this citation style:

-   One author: Tarassow (2019) says ...
-   Two authors: Smeekes and Etienne (2018) say
-   Karanasos et. all (2021) say ...

# Data and Time Series Characteristics

## Data

### Dependent variable: Real GDP Growth

-   FEDFUNDS (Federal Funds Rate): This series is quarterly, obtained
    from FRED, covering the period from July 1954 to February 2024. The
    Federal Funds Rate is typically expressed in percentage per annum,
    indicating the interest rate at which depository institutions trade
    federal funds (balances held at Federal Reserve Banks) with each
    other overnight.

-   Government Spending: This series is quarterly, obtained from FRED,
    covering the period from January 1947 to October 2023. Government
    spending data are usually expressed in currency terms, often in
    billions of USD, representing the total government expenditures
    within a given period.

-   Investment: This series is quarterly, obtained from FRED, covering
    the period from January 1947 to October 2023. Investment data
    typically refer to gross private domestic investment and are
    expressed in currency terms, often in billions of USD, indicating
    the amount of investment in business, residential, and inventory
    investments within the economy.

-   Real GDP (Gross Domestic Product): This series is quarterly,
    obtained from FRED, covering the period from January 2002 to
    October 2023. Real GDP is usually expressed in billions of USD
    (adjusted for inflation), measuring the total economic output of a
    country, accounting for changes in price levels or purchasing power.

-   Personal Consumption Expenditures (PCE): This series is quarterly,
    obtained from FRED, covering the period from January 1959 to
    January 2024. PCE data are typically expressed in billions of USD
    (nominal or real), reflecting the total value of goods and services
    consumed by households.

-   Consumer Sentiment: This series is quarterly, obtained from FRED,
    covering the period from November 1952 to January 2024. Consumer
    Sentiment is measured as an index, which reflects the overall health
    of the economy as perceived by consumers, based on their attitudes
    towards current economic conditions and future expectations.

### Dependent variable: Average Weekly Income

-   Average Weekly Earnings: This series is quarterly, obtained from
    FRED, covering the period from March 2006 to February 2024. The data
    is likely in USD, reflecting average weekly earnings across the
    economy.

-   CPI (Consumer Price Index): The CPI series is quarterly, sourced
    from FRED, covering an extensive period from January 1913 to
    February 2024. It is measured as an index, reflecting changes in the
    price level of a basket of consumer goods and services.

-   Labor Productivity: This series is quarterly, obtained from FRED,
    covering the period from April 1947 to October 2023. The data is
    presented as an index, indicating the efficiency of labor input in
    the production process.

-   Unemployment Rate: The Unemployment Rate series is quarterly,
    obtained from FRED, covering the period from January 1948 to
    February 2024. The data is in percentage terms, representing the
    fraction of the labor force that is not currently employed but
    actively seeking employment.

### Dependent variable: CPI

-   CPI Data: This series represents the Consumer Price Index (CPI) for
    All Urban Consumers (CPI-U), which measures the average change over
    time in the prices paid by urban consumers for a market basket of
    consumer goods and services. The dataset shows monthly data points.
    The data is typically obtained from the Bureau of Labor Statistics
    (BLS). The provided dataset starts from January 1913. However,
    without viewing the entire dataset, the end date is not specified
    here. Generally, CPI data is updated monthly and can run up to the
    most recent full month or the previous month, depending on the
    dataset provided.

-   PPI Data: The Producer Price Index (PPI) series measures the average
    change over time in the selling prices received by domestic
    producers for their output. The data is crucial for understanding
    inflation at the producer level before it impacts consumers.
    Typically sourced from the Bureau of Labor Statistics (BLS).
    Specific dates are not mentioned here, but PPI data is usually
    updated monthly, similar to the CPI.

-   Wage Growth: This dataset likely contains information on the average
    change in wage and salary disbursements over time, which can
    indicate economic health and consumer purchasing power. Data on wage
    growth can be collected from various sources, including the BLS or
    the Bureau of Economic Analysis (BEA), depending on the specific
    measure of wage growth used. As with the other datasets, this
    typically would cover a similar range, often updated on a monthly or
    quarterly basis.

-   Unemployment Rate (UNRATENSA): The unemployment rate data measures
    the percentage of the total labor force that is unemployed but
    actively seeking employment and willing to work. The dataset
    provided seems to focus on non-seasonally adjusted figures. Commonly
    sourced from the Bureau of Labor Statistics (BLS). This dataset is
    usually updated monthly and covers a range that could be similar to
    the other datasets mentioned.

-   Oil Prices: This dataset represents the changes in oil prices over
    time, which can significantly impact global economies and inflation
    rates. Oil prices are a critical component in production costs and
    transportation. Oil price data can be obtained from various sources,
    including the U.S. Energy Information Administration (EIA) or global
    oil market tracking organizations. Oil prices are typically tracked
    daily, but datasets might aggregate this information on a monthly
    basis to align with economic indicators.

### Dependent variable: Real Nonresidential Fixed Investment

-   Interest Rates: BOGZ1FL072052006Q: This dataset likely represents a
    specific economic or financial series, given the code-like name.
    Based on the data structure, it appears to be a quarterly series
    with numerical values associated with each date. Without direct
    reference, the source is not explicitly mentioned, but codes similar
    to this often originate from federal databases, possibly the Board
    of Governors of the Federal Reserve System (U.S. Federal Reserve)
    given the "BOG" prefix. Coverage Period: From July 1954 to at least
    July 1955 (as per the provided sample), indicating a historical
    series that could extend to recent years. The exact end date is not
    provided here, and the series is updated quarterly.

-   Corporate Profits - Q0973BUSQ027NNBR: This dataset also features a
    code-like title, suggesting it's from a specific economic database.
    It contains numerical values recorded quarterly, likely representing
    an economic indicator. The "NNBR" in the name suggests a potential
    connection to the National Bureau of Economic Research (NBER) or
    related datasets. Starting from January 1946, this historical
    dataset provides quarterly data, with the sample showing data up to
    January 1947. The full dataset's range and update frequency would
    need further specifics.

-   Technology Inventories - UITITI: Appears to be a monthly dataset
    with numerical values, potentially related to unemployment, labor,
    or another economic measure based on the naming convention. The
    specific source is not indicated by the name alone, but it may be
    associated with governmental or economic research databases,
    possibly labor or economic statistics. The dataset covers from
    January 1992, providing monthly updates. The sample shown extends to
    May 1992, without the end date of the full dataset specified.

-   Real Nonresidential Fixed Investment (ND000336Q): This quarterly
    dataset measures real non-residential fixed investment, capturing
    business investments in physical assets like buildings (excluding
    residential), machinery, and equipment, adjusted for inflation.
    While the source is not explicitly mentioned, data of this nature is
    often obtained from the Bureau of Economic Analysis (BEA) or similar
    economic statistical agencies. Starting from January 2007, with the
    provided sample extending to January 2008. This type of data is
    typically updated quarterly, reflecting changes in business
    investment behaviors over time.

## Time Series Characteristics

Time series plots:

```{r}
data_long |> 
  ggplot(aes(x = DATE, y = Values, color = Series)) +
  geom_line(show.legend = FALSE) +
  ggtitle("Time Series Plot of Real GDP and predictors") +
  facet_grid(Series ~., scales = "free_y")
```

The time series of Real Gdp and its predictors are shown and after
transforming the series to show growthfrom previous lag, we can observe
that the provided series are all mostly stationary. We only observe a
volatile change for all the series around the time of Covid.

KPSS test results for stationary:

```{r}
data_long_ts |> 
  features(Values, features = list(unitroot_kpss))
```

The KPSS test results indicate that the series is stationary, with a
p-value of 0.1 for all variables, besides Interest Rate, which was 0.09.
This suggests that no differencing might be required to achieve
stationarity, which is essential for accurate forecasting

Seasonality plots:

```{r}
data_long_ts |> 
  gg_subseries(Values)
```

The subseries plot highlights clear seasonal patterns within the data,
for government expenditure and real gdp. suggesting that incorporating
seasonal components into our forecasting model could improve accuracy."
However, for the other variables, it doesn't seem like there is a
seasonal factor in the series.

Autocorrelation properties from correlograms:

```{r}
data_long_ts |> 
  ACF(Values, lag_max = 12) |> 
  autoplot()
```

The ACF plot shows, mostly useful for real gdp, that the current value
is highly correlated with a lagged 4, 8, and 12 reading, as it would
make sense from our understanding that the series is seasonal, so the
data from the same quarter provides high correlation.

```{r}
# Decomposition of Dependent Variable
dcmp <- data_ts|>
  model(stl = STL(real_gdp))

components(dcmp) |>
  autoplot()
```

The STL decomposition separates the time series into trend, seasonal,
and remainder components. The trend component shows a rather stationary
line, except at the 2008 and covid mark from significant world events.
Meanwhile the seasonal component reveals somewhat of s a seasonal trend.
The remainder component, ideally resembling random noise,does for the
most part besides again at the covid part.

Correlations of consumption with the lags of income and sentiment

```{r, include=FALSE}
## Correlations with Real GDP and it's predictors
lag_n <- 12
real_gdp_ir <- data_ts |> 
  CCF(real_gdp, IR, lag_max = lag_n) |> 
  rename(lag = lag, FED_rate = ccf)
real_gdp_inv <- data_ts |> 
  CCF(real_gdp, investment, lag_max = lag_n) |> 
  rename(lag = lag, investment = ccf)
real_gdp_pce <- data_ts |> 
  CCF(real_gdp, pce, lag_max = lag_n) |> 
  rename(lag = lag, PCE = ccf)
real_gdp_govt_exp <- data_ts |> 
  CCF(real_gdp, govt_exp, lag_max = lag_n) |> 
  rename(lag = lag, govt_exp = ccf)
real_gdp_sent <- data_ts |> 
  CCF(real_gdp, sent, lag_max = lag_n) |> 
  rename(lag = lag, sent = ccf)

corrs <- real_gdp_ir |> 
  left_join(real_gdp_pce, by = "lag") |> 
  left_join(real_gdp_inv, by = "lag") |> 
  left_join(real_gdp_govt_exp, by = "lag") |> 
  left_join(real_gdp_sent, by = "lag") |> 
  as_tibble() |> 
  filter(row_number() <= 1+lag_n) |> 
  arrange(desc(lag))
```

```{r}
corrs
```

The CCF between real GDP and the predictors are shown. FedFUNDS interest
rate had very weak correlation coefficients so we decided to drop it
from being used in the models, however the other variables did show some
good correlation at some lags, as we can see. Notably, from investment
showing decent correlation at lag 1,4 and 9.

# Application of Forecasting Methods

Specify training and test sets: I specified the 2021-2023 as test set
and the rest as the training set

```{r}
## specify training and test sets
train <- data_ts |>
  filter(year(DATE) <= 2021)
test <- data_ts |>
  filter(year(DATE) > 2021)

autoplot(train, real_gdp) +
  autolayer(test, real_gdp, color= "red")


```

We specified the test set from 2021 onwards so that we could capture the
covid event, and the training set as anything before.

Dummy Variables We include dummy variables for the year of 2008 and the
first quarter of 2020 in order to capture the financial crisis of 2008
as well as the residual errors from the start of covid. Both were done
in a way to capture the residual errors from those periods.

## Exploring TSLM-D models

In this group I have tried these models:

```{r, echo=TRUE}
#### Exploring TSLM-D models

fit_lm <- train |>
  model(
    lm_1 = TSLM(real_gdp ~ season()  + lag(govt_exp,6) + lag(investment,5) + 
                  dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + 
                  dummy_2020Q1 + dummy_2020Q2),
    lm_2 = TSLM(real_gdp ~ season()  + lag(govt_exp,6) + lag(investment,5) + 
                  lag(pce,7) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + 
                  dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2),
    lm_3 = TSLM(real_gdp ~ season()  + lag(govt_exp,7) + lag(investment,5) + 
                  lag(govt_exp,6) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 
                + dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2),
    lm_4 = TSLM(real_gdp ~ season()  + lag(govt_exp,6) + lag(investment,5) + 
                  lag(pce, 7) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + 
                  dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2),
    lm_5 = TSLM(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1) + 
                  lag(pce, 2) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + 
                  dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2),
    lm_6 = TSLM(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1) + 
                  dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4  + 
                  dummy_2020Q1 + dummy_2020Q2),
    lm_7 = TSLM(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1) + 
                  lag(pce,3) + lag(sent, 2) + dummy_2008Q1 + dummy_2008Q2 + 
                  dummy_2008Q3 + dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2)
  )
```

```{r}
glance(fit_lm) |> 
  arrange(AICc) |> 
  select(.model, AICc)

```

Based on the AICs of all the TSLM models, it seems the TSLM_6 performed
best.

Mathematically, the chosen model is the following: \begin{align*}
y_t &= \beta_0 + \beta_{1,1} g_{t-4} + \beta_{1,1} i_{t-2} + \varepsilon_t
\end{align*} where $\varepsilon_t$ is assumed to be white noise (i.e.
mean zero and serially uncorrelated) and the regression also includes 3
seasonal dummies that are not shown for brevity aswell as dummy
variables for the 2008 year and the first two quarters of 2020. Note
that we ended up chosing $y_{t-1}$ and $s_{t-1}$, which are exactly the
lags where the positive correlation of these two variables with real gdp
is highest as seen in the correlation table in the previous section.

Do residuals diagnostics and try to improve

```{r}
fit_lm |> 
  select(lm_6) |> 
  gg_tsresiduals(lag = 12)
```

The residuals seem to have an average centered at zero which is a good
sign for our model along with the ACF that indicates good significance
of the model. The residuals show that the model has pretty good fit with
the training data but also that residuals seem to be kept in check,
partially with the introduction of dummy variables taking out the events
of covid and the financial crisis of 2008.

## Exploring TSLM-D-ARIMA models

In this group, I will take the model chosen in the TSLM-D group as my
base model and see whether fitting an ARIMA model for the error term is
going to make any difference.

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_lm_arima <- train |>
  model(
    lm_arima_1 = ARIMA(real_gdp ~ season()  + lag(govt_exp,7) + lag(investment,5) 
                       + lag(govt_exp,6)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                       + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_2 = ARIMA(real_gdp ~ 0 + season() + pdq(1,0,2) + PDQ(2,0,1) + 
                         lag(govt_exp,7) + 
                         lag(investment, 5) + lag(govt_exp, 6)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                       + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_3 = ARIMA(real_gdp ~ 0 + season() + pdq(1,0,3) + PDQ(2,0,0) + 
                         lag(govt_exp,7) + 
                         lag(investment, 5) + lag(govt_exp, 6)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                       + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_4 = ARIMA(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1) 
                       + lag(pce, 2)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                       + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_5 = ARIMA(real_gdp ~ season() + 0 + lag(govt_exp,1) + lag(investment,1)
                       + pdq(0:1,0,0:2) + PDQ(0:2,0,0:2)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                       + dummy_2020Q1 + dummy_2020Q2)
  )
```

It seems all specifications for $\eta_t$ lead to the same ARIMA model:
`ARIMA(0,0,1)`

```{r, include=FALSE}
fit_lm_arima |> 
  pivot_longer(everything())
```

Mathematically, `ARIMA(0,0,1)` means `ARIMA()` function suggests the
best model for $\eta_t$ is the following

$$\eta_t = \theta_1 \varepsilon_{t-1} + \varepsilon_t$$ where
$\varepsilon_t$ is a white noise process. Consequently, the full model
becomes $$
y_t = \beta_0 + \beta_{1,1} g_{t-1} + \beta_{1,1} i_{t-1} + \theta_1 \varepsilon_{t-1} + \varepsilon_t 
$$

```{r}
glance(fit_lm_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the TSLM-D-ARIMA models, it seems the
LM_ARIMA_4 performed best. Let's pick `lm_arima_4` from this group.

## ARIMA models

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_arima <- train |> model(
  auto = ARIMA(real_gdp ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
               + dummy_2020Q1 + dummy_2020Q2),  
  auto_s = ARIMA(real_gdp ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                 + dummy_2020Q1 + dummy_2020Q2, stepwise = FALSE),  
  arima_p = ARIMA(real_gdp ~ season() + 0 + pdq(1,0,1) + PDQ(0,0,0)
                  + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + 
                    dummy_2020Q1 + dummy_2020Q2),
  arima002001 = ARIMA(real_gdp ~ season() + 0 + pdq(1,0,2) + PDQ(1,0,1)
                      + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                      + dummy_2020Q1 + dummy_2020Q2),
  arima202202 = ARIMA(real_gdp ~ season() + 0 + pdq(1,0,0) + PDQ(1,0,1)
                      + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                      + dummy_2020Q1 + dummy_2020Q2),
  arima111 = ARIMA(real_gdp ~ season() + 0 + pdq(1,1,1)
                   + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + 
                     dummy_2020Q1 + dummy_2020Q2),
  arima010 = ARIMA(real_gdp ~ season() + 0 + pdq(0,1,0)
                   + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + 
                     dummy_2020Q1 + dummy_2020Q2),
  arima_seasonal = ARIMA(real_gdp ~ season() + 0 + PDQ(1,0,1)
                         + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                         + dummy_2020Q1 + dummy_2020Q2),
  arima_complex = ARIMA(real_gdp ~ season() + 0 + pdq(2,0,2) + PDQ(2,0,2)
                        + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                        + dummy_2020Q1 + dummy_2020Q2),
  arima_stepwise_off = ARIMA(real_gdp ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 
                             + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2, stepwise = FALSE, approximation = FALSE),
  arima_non_seasonal = ARIMA(real_gdp ~ season() + 0 + pdq(0,0,2)
                             + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                             + dummy_2020Q1 + dummy_2020Q2),
  arima_seasonal_adjust = ARIMA(real_gdp ~ season() + 0 + pdq(2,0,2) + PDQ(1,0,1)
                                + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 
                                + dummy_2020Q1 + dummy_2020Q2)
  )  
```

Tabulating the proposed specifications

```{r, include=FALSE}
fit_arima |> 
  pivot_longer(everything(), 
                    names_to = "Model name",
                    values_to = "Orders")
```

Using the Akaike information criterion (AICs) we pick the winning among
ARIMA

```{r}
glance(fit_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the ARIMA models, it seems the auto arima model
performed best.

## Comparing three models on the test set

```{r, echo=TRUE}
fit_train <- train |>
  model(
    lm_1       = TSLM(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1)
                      + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                      + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_1 = ARIMA(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1)
                       + lag(pce, 2)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                       + dummy_2020Q1 + dummy_2020Q2),
    auto       = ARIMA(real_gdp ~ season() + 
                         dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4
                       + dummy_2020Q1 + dummy_2020Q2)
  )
```

Note that I'm manually specifying ARIMA models for `lm_arima_1` and
`auto` as chosen in the previous section. Since we are again estimating
on the same train set, but this would be important if we were to
estimate one of these ARIMA models on the full sample. Because without
these restrictions, ARIMA() function will probably choose another model,
which might not be what we want.

When the three models, one from each group, are compared on the test set
with respect to the RMSE measure, `lm_6` turns out to be the winner

```{r}
fc <- fit_train |> 
  forecast(test)
accuracy(fc, test) |> 
  arrange(RMSE)
```

The AICc values help compare the relative quality of the models while
adjusting for model complexity. Lower AICc values indicate a better fit
to the data, given the number of parameters. This comparison can guide
the selection of the most appropriate model for forecasting real GDP,
balancing fit and parsimony. By comparing all the best models from each
respective model type, we see that the TSLM model performs better with
the lowest RMSE value.

## Obtaining out-of-sample forecasts

Next we estimate the winner model using the full sample

```{r, echo=TRUE, eval=TRUE}
fit_full <- data_ts |>
  model(
    lm_1 = ARIMA(real_gdp ~ season()  + lag(govt_exp,1) + lag(investment,1) + 
                   pdq(0,0,0) + PDQ(0,0,0) 
                 + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 +
                   dummy_2020Q1 + dummy_2020Q2)
  )
```

Due to a bug in the TSLM function, we use the ARIMA model with all p,d,
q, P, D, and Q settings set to 0.

Residual diagnostics and see if you can improve

```{r}
fit_full |> 
  select(lm_1) |> 
  gg_tsresiduals()
```

How do you generate future values for predictors? I used the ARIMA() in
order to generate future values for our predictors based on the
historical data of those predictors and the automatic algorithm from
ARIMA() that will forecast those future values.

```{r, include=FALSE}
gov_exp_fit <- data_ts |>
  model(
    ARIMA(govt_exp)
  )
gov_exp_forecast <- gov_exp_fit |>
  forecast(h = 4)
gov_exp_forecast

inv_fit <- data_ts |>
  model(
    ARIMA(investment)
  )
inv_forecast <- inv_fit |>
  forecast(h = 4)
inv_forecast

gov_exp_forecast_values <- gov_exp_forecast$.mean %>% as.vector()
inv_forecast_values <- inv_forecast$.mean %>% as.vector()


data_future <- new_data(data_ts, 4) |> 
  mutate(
    govt_exp = gov_exp_forecast_values, 
    investment = inv_forecast_values,
    dummy_2008Q1 = 0, 
    dummy_2008Q2 = 0,
    dummy_2008Q3 = 0,
    dummy_2008Q4 = 0,
    dummy_2020Q1 = 0,  
    dummy_2020Q2 = 0
  )
data_future
```

Forecasting next 4 quarters:

```{r}
fc_full <- fit_full |> 
  forecast(new_data = data_future)
fc_full

fc_full |> 
  autoplot(data_ts, level = NULL)
```

#zaz stuff

## Time Series Characteristics

Time series plots:

```{r}
data_long2 |> 
  ggplot(aes(x = DATE, y = Values, color = Series)) +
  geom_line(show.legend = FALSE) +
  facet_grid(Series ~., scales = "free_y")
```

The time series plot of CPI and its predictors reveals no clear trend in
the data over time, iwhile showing to be mostly stationary. While some
vairables show periods of stability, others exhibit significant
fluctuations, which could be considered outliers or unusual
observations. These fluctuations may relate to specific economic events
or changes, underlining the importance of context when interpreting the
data. The overall volatility of the series is around the time of COVID

KPSS test results for stationary:

```{r}
data_long_ts2 |> 
  features(Values, features = list(unitroot_kpss))
```

The KPSS test results indicate that the series is stationary, with a
p-value around or greater than 0.1 for all variables, besides oil
prices, which was .01. This suggests that no differencing might be
required to achieve stationarity, which is essential for accurate
forecasting.

Seasonality plots:

```{r}
data_long_ts2 |> 
  gg_subseries(Values)
```

The subseries plot highlights that the data is non-seasonal for all
variables.

Autocorrelation properties from correlograms:

```{r}
data_long_ts2 |> 
  ACF(Values, lag_max = 12) |> 
  autoplot()
```

The ACF plot shown, proves to be mostly useful for CPI. The current
value is highly correlated with a lagged 4 reading. This makes sense,
given the series is non-seasonal, so the lags are not correlated in a
specific way.

```{r}
# Decomposition of Dependent Variable
dcmp <- data_ts2|>
  model(stl = STL(CPI))

components(dcmp) |>
  autoplot()
```

The STL decomposition separates the time series into trend, seasonal,
and remainder components. The trend component shows a rather stationary
line for all years after 1950, except for subtle bumps in the covid era.
While, the seasonal component reveals some seasonality that varies in
magnitude. The remainder component shows that the series is stationary,
similar to white noise.

Correlations of CPI with the lags of Wage Growth, Oil Prices, PPI,
Exchange Rates, and Unemployment Rates

```{r, include=FALSE}
## Correlations with CPI and it's predictors
lag_n <- 12
CPI_Wagegrowth <- data_ts2 |> 
  CCF(CPI, Wage_Growth, lag_max = lag_n) |> 
  rename(lag = lag, Wage_Growth = ccf)
CPI_OilPrices <- data_ts2 |> 
  CCF(CPI, Oil_Prices, lag_max = lag_n) |> 
  rename(lag = lag, Oil_Prices = ccf)
CPI_PPI <- data_ts2 |> 
  CCF(CPI, PPI, lag_max = lag_n) |> 
  rename(lag = lag, PPI = ccf)
CPI_ExchangeRates <- data_ts2 |> 
  CCF(CPI, Exchange_Rates, lag_max = lag_n) |> 
  rename(lag = lag, Exchange_Rates = ccf)
CPI_UnempRates <- data_ts2 |> 
  CCF(CPI, Unemp_Rate, lag_max = lag_n) |> 
  rename(lag = lag, Unemp_Rate = ccf)

corrs <- CPI_Wagegrowth |> 
  left_join(CPI_OilPrices, by = "lag") |>
  left_join(CPI_PPI, by = "lag") |>
  left_join(CPI_ExchangeRates, by = "lag") |>
  left_join(CPI_UnempRates, by = "lag") |> 
  as_tibble() |> 
  filter(row_number() <= 1+lag_n) |> 
  arrange(desc(lag))
```

```{r}
corrs
```

The CCF between CPI and the predictors are shown. Wage growth and
exchange rates had very weak correlation coefficients so we decided to
drop it from being used in the models, however the other variables did
show some good correlation at some lags, as we can see. Specifically,
from Oil Prices and PPI showing great correlation at lag 1 and 2.

# Application of Forecasting Methods

Specify training and test sets: I specified the 2021-2023 as test set
and the rest as the training set

```{r}
## specify training and test sets
train <- data_ts2 |>
  filter(year(DATE) <= 2021)
test <- data_ts2 |>
  filter(year(DATE) > 2021)

autoplot(train, CPI) +
  autolayer(test, CPI, color= "red")


```

We specified the test set from 2021 onwards so that we could capture the
covid event, and the training set as anything before.

Dummy Variables We include dummy variables for the year of 2008 to
capture the financial crisis of 2008. This was done in a way to capture
the residual errors from that period.

## Exploring TSLM-D models

In this group I have tried these models:

```{r, echo=TRUE}
#### Exploring TSLM-D models

fit_lm <- train |> 
  model(
    lm_1 = TSLM(CPI ~ season() + lag(Oil_Prices,2) + lag(PPI,1) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
    lm_2 = TSLM(CPI ~ season() + lag(Oil_Prices,2) + lag(PPI,1) + lag(Oil_Prices,3) + lag(PPI, 2)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
    lm_3 = TSLM(CPI ~ season() + lag(Oil_Prices, 1) + lag(PPI, 3)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
    lm_4 = TSLM(CPI ~ season() + lag(Oil_Prices,1) + lag(Oil_Prices, 2) + lag(PPI, 1) + lag(PPI, 2)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4)
  )
```

```{r}
glance(fit_lm) |> 
  arrange(AICc) |> 
  select(.model, AICc)

```

Based on the AICs of all the TSLM models, it seems that TSLM4 performed
the best, though TSLM2 had a lower RMSE, so we came to the conclusion to
use TSLM2.

Mathematically, the chosen model is the following: \begin{align*}
y_t &= \beta_0 + \beta_{1,1} g_{t-4} + \beta_{1,1} i_{t-2} + \varepsilon_t
\end{align*} where $\varepsilon_t$ is assumed to be white noise (i.e.
mean zero and serially uncorrelated) and the regression also includes 3
seasonal dummies that are not shown for brevity aswell as dummy
variables for the 2008 year and the first two quarters of 2020. Note
that we ended up chosing $y_{t-1}$ and $s_{t-1}$, which are exactly the
lags where the positive correlation of these two variables with real gdp
is highest as seen in the correlation table in the previous section.
KURTIK \*\* DO THIS FOR CPI

Do residuals diagnostics and try to improve

```{r}
fit_lm |> 
  select(lm_2) |> 
  gg_tsresiduals(lag = 12)
```

The residuals seem to have an average centered at zero which is a good
sign for our model along with the ACF that indicates good significance
of the model. The residuals show that the model has pretty good fit with
the training data but also that residuals seem to be kept in check,
partially with the introduction of dummy variables taking out the
financial crisis of 2008.

## Exploring TSLM-D-ARIMA models

In this group, I will take the model chosen in the TSLM-D group as my
base model and see whether fitting an ARIMA model for the error term is
going to make any difference.

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_lm_arima <- train |>
  model(
    lm_arima_1 = ARIMA(CPI ~ season() + lag(Oil_Prices,2) + lag(PPI,1)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
    lm_arima_2 = ARIMA(CPI ~ season() + lag(Oil_Prices,2) + lag(PPI,1) 
                       + pdq(0:2, 0, 0:2) + PDQ(0:1, 0, 0) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4)
  )
```

```{r, include=FALSE}
fit_lm_arima |> 
  pivot_longer(everything())
```

Mathematically, `ARIMA(0,0,1)` means `ARIMA()` function suggests the
best model for $\eta_t$ is the following

$$\eta_t = \theta_1 \varepsilon_{t-1} + \varepsilon_t$$ where
$\varepsilon_t$ is a white noise process. Consequently, the full model
becomes $$
y_t = \beta_0 + \beta_{1,1} g_{t-1} + \beta_{1,1} i_{t-1} + \theta_1 \varepsilon_{t-1} + \varepsilon_t 
$$ KURTIK DO THIS \*\*

```{r}
glance(fit_lm_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the TSLM-D-ARIMA models, it seems the
LM_ARIMA_2 performed best, due to the AICc being lower.

## ARIMA models

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_arima <- train |> model(
  auto = ARIMA(CPI ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),  
  auto_s = ARIMA(CPI ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4, stepwise = FALSE),  
  arima_p = ARIMA(CPI ~ season() + 0 + pdq(0,0,0) + PDQ(0:3,1,0:3) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
  arima002001 = ARIMA(CPI ~ season() + 0 + pdq(0,0,2) + PDQ(0,0,1) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
  arima202202 = ARIMA(CPI ~ season() + 0 + pdq(2,0,0) + PDQ(0,0,1) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4))  
```

Tabulating the proposed specifications

```{r, include=FALSE}
fit_arima |> 
  pivot_longer(everything(), 
               names_to = "Model name",
               values_to = "Orders")
```

Using the Akaike information criterion (AICs) we pick the winning among
ARIMA

```{r}
glance(fit_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the -ARIMA models, it seems the auto arima
model performed best.

## Comparing three models on the test set

```{r, echo=TRUE}
fit_train <- train |>
  model(
    lm_1       = TSLM(CPI ~ season() + 0 + lag(Oil_Prices,2) + lag(PPI,1) + lag(Oil_Prices,3) + lag(PPI, 2)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
    lm_arima_2 = ARIMA(CPI ~ season() + 0 + lag(Oil_Prices,2) + lag(PPI,1) 
                       + pdq(0:2, 0, 0:2) + PDQ(0:1, 0, 0)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4),
    auto       = ARIMA(CPI ~ season() + 0 + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4)
  )
```

Note that I'm manually specifying ARIMA models for `lm_arima_1` and
`auto` as chosen in the previous section. Since we are again estimating
on the same train set, but this would be important if we were to
estimate one of these ARIMA models on the full sample. Because without
these restrictions, ARIMA() function will probably choose another model,
which might not be what we want.

When the three models, one from each group, are compared on the test set
with respect to the RMSE measure, `lm_6` turns out to be the winner

```{r}
fc <- fit_train |> 
  forecast(test)
accuracy(fc, test) |> 
  arrange(RMSE)
```

The AICc values help compare the relative quality of the models while
adjusting for model complexity. Lower AICc values indicate a better fit
to the data, given the number of parameters. This comparison can guide
the selection of the most appropriate model for forecasting CPI,
balancing fit and parsimony. By comparing all the best models from each
respective model type, we see that the TSLM model performs better with
the lowest RMSE value.

## Obtaining out-of-sample forecasts

Next we estimate the winner model using the full sample

```{r, echo=TRUE, eval=TRUE}
fit_full <- data_ts2 |>
  model(
    lm_1 = ARIMA(CPI ~ season() + 0 + lag(Oil_Prices,2) + lag(PPI,1) + lag(Oil_Prices,3) + lag(PPI, 2) 
                 + pdq(0, 0, 0) + PDQ(0, 0, 0)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4))
```

Due to a bug in the TSLM function, we use the ARIMA model with all p,d,
q, P, D, and Q settings set to 0.

Residual diagnostics and see if you can improve model further:

```{r}
fit_full |> 
  select(lm_1) |> 
  gg_tsresiduals()
```

How do you generate future values for predictors? I used the ARIMA() in
order to generate future values for our predictors based on the
historical data of those predictors and the automatic algorithm from
ARIMA() that will forecast those future values.

```{r, include=FALSE}
Oil_Prices_fit <- data_ts2 |>
  model(
    ARIMA(Oil_Prices)
  )
Oil_Prices_forecast <- Oil_Prices_fit |>
  forecast(h = "12 months")
Oil_Prices_forecast


PPI_fit <- data_ts2 |>
  model(
    ARIMA(PPI)
  )
PPI_forecast <- PPI_fit |>
  forecast(h = "12 months")
PPI_forecast

Oil_Prices_Forecast_values <- Oil_Prices_forecast$.mean |> as.vector()
PPI_Forecast_values <- PPI_forecast$.mean |> as.vector()


data_future <- new_data(data_ts2, 12) |> 
  mutate(
    Oil_Prices = Oil_Prices_Forecast_values, 
    PPI = PPI_Forecast_values,
    dummy_2008Q1 = 0, 
    dummy_2008Q2 = 0,
    dummy_2008Q3 = 0,
    dummy_2008Q4 = 0,
    dummy_2020Q1 = 0,  
    dummy_2020Q2 = 0
  )
data_future
```

Forecasting next 12 months:

```{r}
fc_full <- fit_full |> 
  forecast(new_data = data_future)
fc_full

fc_full |> 
  autoplot(data_ts2, level = NULL)
```

#RNFI

## Time Series Characteristics

Time series plots:

```{r}
data_long3 |> 
  ggplot(aes(x = DATE, y = Values, color = Series)) +
  geom_line(show.legend = FALSE) +
  facet_grid(Series ~., scales = "free_y")
```

The time series plot of Real Nonresidential Fixed Investment and its
predictors shows that the series is mostly stationary, except for the
time period of COVID. To further analyze this, a KPSS needs to be ran to
identify if the variables are stationary or not.

KPSS test results for stationary:

```{r}
data_long_ts3 |> 
  features(Values, features = list(unitroot_kpss))
```

The KPSS test results indicate that the series is stationary, with a
p-value around or greater than 0.1 for all variables, besides Real
Estate Values, which was .01. This suggests that no differencing might
be required to achieve stationarity, which is essential for accurate
forecasting.

Seasonality plots:

```{r}
data_long_ts3 |> 
  gg_subseries(Values)
```

The subseries plot highlights clear seasonal patterns within the data,
for RNFI and Tech_Inv, though not for Interest Rates and Real estate
values. These patterns suggest that incorporating seasonal components
into our forecasting model could improve accuracy for the two variables
that had seasonality.

Autocorrelation properties from correlograms:

```{r}
data_long_ts3 |> 
  ACF(Values, lag_max = 12) |> 
  autoplot()
```

The ACF plot shown, proves to be mostly useful for RNFI. The current
value is highly correlated with a lagged 4, 8, and 12 reading. This
makes sense, given the series seems to be seasonal, so the data from the
same quarter provides high correlation.

```{r}
# Decomposition of Dependent Variable
dcmp <- data_ts3 |>
  model(stl = STL(RNFI))

components(dcmp) |>
  autoplot()
```

The STL decomposition separates the time series into trend, seasonal,
and remainder components. The trend component shows no clear trend
throughout the years. While, the seasonal component reveals a ton of
seasonality. The remainder component shows that the series is
stationary, except for the period of COVID, similar to white noise.

Correlations of RNFI with the lags of Interest Rates, Tech Inventories,
and Real Estate Market Value

```{r, include=FALSE}
## Correlations with RNFI and it's predictors
lag_n <- 12
RNFI_IntRates <- data_ts3 |> 
  CCF(RNFI, Int_Rates, lag_max = lag_n) |> 
  rename(lag = lag, Int_Rates = ccf)
RNFI_TechInv <- data_ts3 |> 
  CCF(RNFI, Tech_Inv, lag_max = lag_n) |> 
  rename(lag = lag, Tech_Inv = ccf)
RNFI_RealEstateMV <- data_ts3 |> 
  CCF(RNFI, Real_Estate_MV, lag_max = lag_n) |> 
  rename(lag = lag, Real_Estate_MV = ccf)

corrs <- RNFI_IntRates |> 
  left_join(RNFI_TechInv, by = "lag") |>
  left_join(RNFI_RealEstateMV, by = "lag") |> 
  as_tibble() |> 
  filter(row_number() <= 1+lag_n) |> 
  arrange(desc(lag))

```

```{r}
corrs
```

The CCF between RNFI and the predictors are shown. Interest Rates had
very weak correlation coefficients so we decided to drop it from being
used in the models, however the other variables did show some good
correlation at some lags, as we can see. Specifically, from Tech
Inventory and Real Estate Values showing great correlation at lag 1 and
5.

# Application of Forecasting Methods

Specify training and test sets: I specified the 2021-2023 as test set
and the rest as the training set

```{r}
## specify training and test sets
train <- data_ts3 |>
  filter(year(DATE) <= 2021)
test <- data_ts3 |>
  filter(year(DATE) > 2021)

autoplot(train, RNFI) +
  autolayer(test, RNFI, color= "red")


```

We specified the test set from 2021 onwards so that we could capture the
covid event, and the training set as anything before.

Dummy Variables We include dummy variables for the year of 2008 to
capture the financial crisis of 2008 and the first quarter of 2020 to
capture the residual errors from the start of COVID. This was done in a
way to capture the residual errors from those periods.

## Exploring TSLM-D models

In this group I have tried these models:

```{r, echo=TRUE}
#### Exploring TSLM-D models

fit_lm <- train |>
  model(
    lm_1 = TSLM(RNFI ~ season() + lag(Tech_Inv,1) + lag(Real_Estate_MV,1) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_2 = TSLM(RNFI ~ season() + lag(Tech_Inv,5) + lag(Real_Estate_MV,2) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_3 = TSLM(RNFI ~ season() + lag(Tech_Inv,1) + lag(Tech_Inv,5) + lag(Real_Estate_MV,1) + lag(Real_Estate_MV, 5) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_4 = TSLM(RNFI ~ season() + lag(Tech_Inv,1) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_5 = TSLM(RNFI ~ season() + lag(Tech_Inv, 1) + lag(Tech_Inv,5) + lag(Tech_Inv,9)+ lag(Real_Estate_MV,1) + lag(Real_Estate_MV, 2) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2))
```

```{r}
glance(fit_lm) |> 
  arrange(AICc) |> 
  select(.model, AICc)

```

Based on the AICs of all the TSLM models, it seems that TSLM5 performed
the best

Mathematically, the chosen model is the following: \begin{align*}
y_t &= \beta_0 + \beta_{1,1} g_{t-4} + \beta_{1,1} i_{t-2} + \varepsilon_t
\end{align*} where $\varepsilon_t$ is assumed to be white noise (i.e.
mean zero and serially uncorrelated) and the regression also includes 3
seasonal dummies that are not shown for brevity aswell as dummy
variables for the 2008 year and the first two quarters of 2020. Note
that we ended up chosing $y_{t-1}$ and $s_{t-1}$, which are exactly the
lags where the positive correlation of these two variables with real gdp
is highest as seen in the correlation table in the previous section.
KURTIK \*\* DO THIS FOR RNFI

Do residuals diagnostics and try to improve

```{r}
fit_lm |> 
  select(lm_5) |> 
  gg_tsresiduals(lag = 12)
```

The residual analysis reveals the average is centered around zero,
though there was a lot of randomness that was shown. Overall, the
residuals show that the model has a good fit with the training data,
which could be due to the introduction of dummy variables during 2008
and 2020.

## Exploring TSLM-D-ARIMA models

In this group, I will take the model chosen in the TSLM-D group as my
base model and see whether fitting an ARIMA model for the error term is
going to make any difference.

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_lm_arima <- train |>
  model(
    lm_arima_1 = ARIMA(RNFI ~ season() + 0 + lag(Tech_Inv,1) + lag(Real_Estate_MV,1) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_2 = ARIMA(RNFI ~ season() + 0 + lag(Tech_Inv,5) + lag(Real_Estate_MV,2) + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_3 = ARIMA(RNFI ~ season() + 0 + lag(Tech_Inv,5) + lag(Real_Estate_MV,2) 
                       + pdq(1, 0, 3) + PDQ(2, 0, 0)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_4 = ARIMA(RNFI ~ season() + 0 + lag(Tech_Inv,1) + lag(Real_Estate_MV,1)
                       + pdq(1, 0, 2) + PDQ(2, 0, 1)+ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2))
```

```{r, include=FALSE}
fit_lm_arima |> 
  pivot_longer(everything())
```

Mathematically, `ARIMA(0,0,1)` means `ARIMA()` function suggests the
best model for $\eta_t$ is the following

$$\eta_t = \theta_1 \varepsilon_{t-1} + \varepsilon_t$$ where
$\varepsilon_t$ is a white noise process. Consequently, the full model
becomes $$
y_t = \beta_0 + \beta_{1,1} g_{t-1} + \beta_{1,1} i_{t-1} + \theta_1 \varepsilon_{t-1} + \varepsilon_t 
$$ KURTIK DO THIS \*\*

```{r}
glance(fit_lm_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the TSLM-D-ARIMA models, it seems the
LM_ARIMA_2 performed best, due to the AICc being lower.

## ARIMA models

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_arima <- train |> model(
  auto = ARIMA(RNFI ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2),  
  auto_s = ARIMA(RNFI ~ dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2, stepwise = FALSE),  
  arima_p = ARIMA(RNFI ~ season() + 0 + pdq(1,0,1) + PDQ(0,0,0)
                  + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2),
  arima002001 = ARIMA(RNFI ~ season() + 0 + pdq(1,0,2) + PDQ(1,0,1)
                      + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2),
  arima202202 = ARIMA(RNFI ~ season() + 0 + pdq(1,0,0) + PDQ(1,0,1)
                      + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2))  
```

Tabulating the proposed specifications

```{r, include=FALSE}
fit_arima |> 
  pivot_longer(everything(), 
               names_to = "Model name",
               values_to = "Orders")
```

Using the Akaike information criterion (AICs) we pick the winning among
ARIMA

```{r}
glance(fit_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the -ARIMA models, it seems the auto arima
model performed best.

## Comparing three models on the test set

```{r, echo=TRUE}
fit_train <- train |>
  model(
    lm_1       = TSLM(RNFI ~ season() + 0 + lag(Tech_Inv,1) + lag(Real_Estate_MV,1) + lag(Tech_Inv, 5) + lag(Real_Estate_MV, 5) 
                      + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4 + dummy_2020Q1 + dummy_2020Q2),
    lm_arima_1 = ARIMA(RNFI ~ season() + 0 + lag(Tech_Inv,1) + lag(Real_Estate_MV,1) + lag(Tech_Inv, 5)
                       + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2),
    auto       = ARIMA(RNFI ~ season() + 0 +  
                       dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4  + dummy_2020Q1 + dummy_2020Q2)
  )
```

Note that I'm manually specifying ARIMA models for `lm_arima_1` and
`auto` as chosen in the previous section. Since we are again estimating
on the same train set, but this would be important if we were to
estimate one of these ARIMA models on the full sample. Because without
these restrictions, ARIMA() function will probably choose another model,
which might not be what we want.

\*\*\*IS THIS THE SAME? \^ and below

When the three models, one from each group, are compared on the test set
with respect to the RMSE measure, `lm_6` turns out to be the winner

```{r}
fc <- fit_train |> 
  forecast(test)
accuracy(fc, test) |> 
  arrange(RMSE)
```

The AICc values help compare the relative quality of the models while
adjusting for model complexity. Lower AICc values indicate a better fit
to the data, given the number of parameters. This comparison can guide
the selection of the most appropriate model for forecasting RNFI,
balancing fit and parsimony. By comparing all the best models from each
respective model type, we see that the TSLM-D-ARIMA model performs
better with the lowest RMSE value.

## Obtaining out-of-sample forecasts

Next we estimate the winner model using the full sample

```{r, echo=TRUE, eval=TRUE}
fit_full <- data_ts3 |>
  model(
    lm_arima_1 = ARIMA(RNFI ~ season()  + lag(Tech_Inv,1) + lag(Real_Estate_MV,1) + lag(Tech_Inv, 5) + 
                   pdq(0,0,0) + PDQ(0,0,0) 
                 + dummy_2008Q1 + dummy_2008Q2 + dummy_2008Q3 + dummy_2008Q4   + dummy_2020Q1 + dummy_2020Q2)
  )
```

Due to a bug in the TSLM function, we use the ARIMA model with all p,d,
q, P, D, and Q settings set to 0.

Residual diagnostics and see if you can improve model further:

```{r}
fit_full |> 
  select(lm_arima_1) |> 
  gg_tsresiduals()
```

How do you generate future values for predictors? I used the ARIMA() in
order to generate future values for our predictors based on the
historical data of those predictors and the automatic algorithm from
ARIMA() that will forecast those future values. \*\*\*\*\*\*

```{r, include=FALSE}
RealEstate_MV_fit <- data_ts3 |>
  model(
    ARIMA(Real_Estate_MV)
  )
RealEstate_MV_forecast <- RealEstate_MV_fit |>
  forecast(h = 4)
RealEstate_MV_forecast

TechInv_fit <- data_ts3 |>
  model(
    ARIMA(Tech_Inv)
  )
TechInv_forecast <- TechInv_fit |>
  forecast(h = 4)
TechInv_forecast


RealEstate_MV_forecast_values <- RealEstate_MV_forecast$.mean |>  as.vector()
TechInv_forecast_values <- TechInv_forecast$.mean |> as.vector()


data_future <- new_data(data_ts3, 4) |> 
  mutate(
    Real_Estate_MV = RealEstate_MV_forecast_values, 
    Tech_Inv = TechInv_forecast_values,
    dummy_2008Q1 = 0, 
    dummy_2008Q2 = 0,
    dummy_2008Q3 = 0,
    dummy_2008Q4 = 0,
    dummy_2020Q1 = 0,  
    dummy_2020Q2 = 0
  )
data_future
```

Forecasting next 4 quarters:

```{r}
fc_full <- fit_full |> 
  forecast(new_data = data_future)

fc_full

fc_full |> 
  autoplot(data_ts3)
```

# Average Weekly Earning

## Time Series Characteristics

Time series plots:

```{r}
data_long4 |> 
  ggplot(aes(x = DATE, y = Values, color = Series)) +
  geom_line(show.legend = FALSE) +
  ggtitle("Time Series Plot of Avg Weekly Earnings and predictors") +
  facet_grid(Series ~., scales = "free_y")
```

The time series of average weekly earnings and its predictors are shown and after transforming the series to show growth from previous lag, we can observe that the provided series are all mostly stationary. We only observe a volatile change for all the series around the time of Covid.

KPSS test results for stationary:

```{r}
data_long_ts4 |> 
  features(Values, features = list(unitroot_kpss))
```

The KPSS test results indicate that the series is stationary, with a test statistic of 0.1 for all variables. This suggests that no differencing might be required to achieve stationarity, which is essential for accurate forecasting.

Seasonality plots:

```{r}
data_long_ts4 |> 
  gg_subseries(Values)
```

The subseries plot highlights clear seasonal patterns within the data,
for cpi and average weekly earnings. suggesting that incorporating
seasonal components into our forecasting model could improve accuracy."
However, for the other variables, it doesnt seem like there is a
seasonal factor in the series.

Autocorrelation properties from correlograms:

```{r}
data_long_ts4 |> 
  ACF(Values, lag_max = 12) |> 
  autoplot()
```

The ACF plot shows, mostly useful for average weekly earnings, that the current value is highly correlated with a lagged 3,6 and 10 reading, as it would make sense from our understanding that the series is seasonal.

```{r}
dcmp <- data_ts4 |>
  model(stl = STL(avg_weekly_earnings))

components(dcmp) |>
  autoplot()
```

The STL decomposition separates the time series into trend, seasonal,
and remainder components. The trend component shows a rather stationary
line, except at the 2008 and covid mark from significant world events.
Meanwhile the seasonal component reveals somewhat of a seasonal trend,
that increases in magnitude in most recent years. The remainder component, ideally resembling random noise,does for the most part.

Correlations of consumption with the lags of income and sentiment

```{r, include=FALSE}
## Correlations with income and sentiment
lag_n <- 24
awe_cpi <- data_ts4 |> 
  CCF(avg_weekly_earnings, CPI_index, lag_max = lag_n) |> 
  rename(lag = lag, CPI = ccf)
awe_unemp <- data_ts4 |> 
  CCF(avg_weekly_earnings, unemp_rate, lag_max = lag_n) |> 
  rename(lag = lag, Unemp_rate = ccf)
awe_emp <- data_ts4 |> 
  CCF(avg_weekly_earnings, emp_rate, lag_max = lag_n) |> 
  rename(lag = lag, Emp_rate = ccf)
awe_ind_index <- data_ts4 |> 
  CCF(avg_weekly_earnings, index, lag_max = lag_n) |> 
  rename(lag = lag, Ind_index = ccf)

corrs <- awe_cpi |> 
  left_join(awe_unemp, by = "lag") |> 
  left_join(awe_emp, by = "lag") |> 
  left_join(awe_ind_index, by = "lag") |> 
  as_tibble() |> 
  filter(row_number() <= 1+lag_n) |> 
  arrange(desc(lag))
```

```{r}
corrs
```

The CCF between average weekly earnings and the predictors are shown. Most predictors had weak correlation across the board, indicating that perhaps an arima model using lagged instances of the dependent variable might perform best. The best correlations came from a newly introduced predictor for industry index.

# Application of Forecasting Methods

Specify training and test sets: I specified the 2021-2023 as test set
and the rest as the training set

```{r}
## specify training and test sets
train <- data_ts4 |>
  filter(year(DATE) <= 2021)
test <- data_ts4 |>
  filter(year(DATE) > 2021)

autoplot(train, avg_weekly_earnings) +
  autolayer(test, avg_weekly_earnings, color= "red")


```

We specified the test set from 2021 onwards so that we could capture the
covid event, and the training set as anything before.

## Exploring TSLM-D models

In this group I have tried these models:

```{r, echo=TRUE}
#### Exploring TSLM-D models

fit_lm <- train |>
  model(
    lm_1 = TSLM(avg_weekly_earnings ~ season() + lag(CPI_index,5) + lag(emp_rate,4) + lag(unemp_rate,4)),
    lm_2 = TSLM(avg_weekly_earnings ~ season() + lag(CPI_index,5) + lag(CPI_index,6) + lag(unemp_rate, 4) + lag(emp_rate,4) + lag(emp_rate,7)),
    lm_3 = TSLM(avg_weekly_earnings ~ season() + lag(CPI_index,5)),
    lm_4 = TSLM(avg_weekly_earnings ~ season() + lag(index, 9) + lag(CPI_index,5) + lag(CPI_index,2)),
    lm_5 = TSLM(avg_weekly_earnings ~ season() + lag(unemp_rate, 4)),
    lm_6 = TSLM(avg_weekly_earnings ~ season() + lag(index, 9) + lag(CPI_index,5))
  )
```

```{r}
glance(fit_lm) |> 
  arrange(AICc) |> 
  select(.model, AICc)

```

Based on the AICs of all the TSLM models, it seems the TSLM6 performed best, using only a lagged instance of CPI and industry index. 

Mathematically, the chosen model is the following: \begin{align*}
y_t &= \beta_0 + \beta_{1,1} i_{t-9} + \beta_{1,1} c_{t-5} + \varepsilon_t
\end{align*} where $\varepsilon_t$ is assumed to be white noise (i.e.
mean zero and serially uncorrelated) and the regression also includes 3
seasonal dummies that are not shown for brevity . 

Do residuals diagnostics and try to improve

```{r}
fit_lm |> 
  select(lm_6) |> 
  gg_tsresiduals(lag = 12)
```
Our residual analysis suggests that our residuals are situated at mean zero so that is a good sign. However we then notice a huge autocorrelation in the ACF graph which isnt very good when justifying our forecast. 

## Exploring TSLM-D-ARIMA models

In this group, I will take the model chosen in the TSLM-D group as my
base model and see whether fitting an ARIMA model for the error term is
going to make any difference.

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_lm_arima <- train |>
  model(
    lm_arima_1 = ARIMA(avg_weekly_earnings ~ season() + 0 + lag(CPI_index,5) + lag(emp_rate,4) + lag(unemp_rate,4) +
                         pdq(0:3,0,0:2) + PDQ(0:1,0,0:1)),
    lm_arima_2 = ARIMA(avg_weekly_earnings ~ season() + 0 + lag(CPI_index,5) + lag(emp_rate,4) + lag(unemp_rate,4), 
                       stepwise = FALSE),
    lm_arima_3 = ARIMA(avg_weekly_earnings ~ season() + 0 + lag(CPI_index,5) + lag(emp_rate,4) + lag(unemp_rate,4) +
                         pdq(2:3,0,0:1) + PDQ(0:2,0,0:1)),
    lm_arima_4 = ARIMA(avg_weekly_earnings ~ season() + lag(index, 9) + lag(CPI_index,5))
  )
```



```{r, include=FALSE}
fit_lm_arima |> 
  pivot_longer(everything())
```


```{r}
glance(fit_lm_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the TSLM-D-ARIMA models, it seems the LM_ARIMA_2 performed best. 

## ARIMA models

In this group I have tried the following specifications:

```{r, echo=TRUE}
fit_arima <- train |> model(
  auto = ARIMA(avg_weekly_earnings ~ season() + 0 ),  
  auto_s = ARIMA(avg_weekly_earnings ~ season() + 0 , stepwise = FALSE), 
  arima_p = ARIMA(avg_weekly_earnings ~ season() + 0 + pdq(0,0,0) + PDQ(0:3,0,0:2)),
  arima1 = ARIMA(avg_weekly_earnings ~ season() + 0 + pdq(0:3,0,0:1) + PDQ(0:1,0,0:1)),
  arima2 = ARIMA(avg_weekly_earnings ~ season() + 0 + pdq(0:2,0,0:2) + PDQ(0:1,0,0:1)),
  arima3 = ARIMA(avg_weekly_earnings ~ season() + 0  +
                       pdq(2:3,0,0:1) + PDQ(0:2,0,0:1))
  )  
```

Tabulating the proposed specifications

```{r, include=FALSE}
fit_arima |> 
  pivot_longer(everything(), 
                    names_to = "Model name",
                    values_to = "Orders")
```

Using the Akaike information criterion (AICs) we pick the winning among
ARIMA

```{r}
glance(fit_arima) |> 
  arrange(AICc) |> 
  select(.model, AICc)
```

Based on the AICs of all the -ARIMA models, it seems the auto_s arima model performed best. 

## Comparing three models on the test set

```{r, echo=TRUE}
fit_train <- train |>
  model(
    lm_1     =  TSLM(avg_weekly_earnings ~ season() + lag(index, 9) + lag(CPI_index,5)),
    lm_arima_1 = ARIMA(avg_weekly_earnings ~ season() + lag(index, 9) + lag(CPI_index,5)),
    auto_s = ARIMA(avg_weekly_earnings ~ season() + 0 , stepwise = FALSE)
  )
```


When the three models, one from each group, are compared on the test set
with respect to the RMSE measure, `lm_6` turns out to be the winner

```{r}
fc <- fit_train |> 
  forecast(test)
accuracy(fc, test) |> 
  arrange(RMSE)
```

The AICc values help compare the relative quality of the models while adjusting for model complexity. Lower AICc values indicate a better fit to the data, given the number of parameters. This comparison can guide the selection of the most appropriate model for forecasting average weekly earnings, balancing fit and parsimony. By comparing all the best models from each respective model type, we see that the auto_s model performs better with the lowest RMSE value.

## Obtaining out-of-sample forecasts

Next we estimate the winner model using the full sample

```{r, echo=TRUE, eval=TRUE}
fit_full <- data_ts4 |>
  model(
    auto_s = ARIMA(avg_weekly_earnings ~ season() + 0 , stepwise = FALSE)
  )
```

Due to a bug in the TSLM function, we use the ARIMA model with all p,d,
q, P, D, and Q settings set to 0.

Residual diagnostics and see if you can improve model further:

```{r}
fit_full |> 
  select(auto_s) |> 
  gg_tsresiduals()
```


Forecasting next 12 months for average weely earnings:

```{r}
fc_full <- fit_full |> 
  forecast(h = 12)

fc_full

fc_full |> 
  autoplot(data_ts4, level = NULL)
```

The final graph visualizes forecasts for future periods based on the selected model, using historical data on the dependent variable, we were able to forecast for the next 12 months as observed on our graph.

# Conclusion
# Limitations and Conclusion

## Limitations

When forecasting Real GDP, we've come to face some limitations in our
methodology and execution methods. These limitations arose from either
the choice of series in terms of variable, historical data,as well as
how they correlate with the dependent variable, in this case Real GDP.
This was the case for CPI as well. One of the challenges was the choice
of models. For most, we use the AIC criteria as means to compare similar
structured models and pick the winning one in that category. However,
when analyzing the accuracy of those models, we came to notice that it
wasn't always the case that the winning model had the lowest RMSE
value. We use AIC and RMSE as a means to compare and pick models, but
when their results seem to oppose, we would tend to pick the model based
on AIC, due to its relative superiority. Perhaps this wasn't the best
choice in hindsight, so that would definitely be something to look into
for future forecasting projects. Another issue we came across in
forecasting our models was dealing with major world events that affected
the series in an unpredictable manner. Events like the 2008 financial
crisis or COVID, gave our series very sudden and unpredictable spikes at
some time intervals which made our training set not entirely
predictable. To deal with those complications, we included dummy
variables to capture the events in 2008 and early 2020. However, just by
viewing the time series, it was hard to get a full grasp on all major
events that would have affected the series and therefore a lot was not
captured in dummy variables and just left in the series. Additionally,
adding too many dummy variables would've limited the data we were
working with so that was another complication. When working with CPI and
RNFI, we realized that some of the predictors were not useful indicators
for forecasting. Specifically, when we checked for correlation for CPI
predictors, the lag values were not greater than 0.1 for Wage Growth,
Unemployment Rates, and Exchange Rates. Because of these results, we
decided it was best to stick with PPI and Oil Prices as predictors,
which had great correlation among different lags. When checking for
correlation between RNFI and its predictors, we came across the same
problem, and decided it was best to exclude Interest Rates in the future
models. Overall, the decision of leaving out these predictors was made
in order to successfully forecast the dependent variables in the best
way possible, avoiding fluctuations in data, and decreasing the
complexity of the models, leading to a lower AICc.

## Conclusion
Forecasting key economic indicatorsReal GDP, CPI, Average Weekly Earnings, and Real Nonresidential Fixed Investmentis crucial for shaping economic policies and investment decisions. This complex task hinges on analyzing vast amounts of data and understanding the interplay between various economic forces.
We employed a range of methodologies, including time series analysis and econometric modeling to predict future economic conditions. In the end, forecasting remains challenging due to the unpredictable nature of economic shocks, such as geopolitical events or pandemics.
Economic forecasting is indispensable for navigating the future economic landscape. It requires a balance of technical expertise and an understanding of global dynamics, with the aim of supporting informed decision-making and fostering a stable economic environment.



```
